---
title: "Extended Mapping Presentation"
author: "Ruxin Liu, Jinzhe Zhang"
date: "11/12/2020"
output:
  revealjs::revealjs_presentation:
    theme: beige
    highlight: pygments
    css: reveal.css
    center: yes
    transition: slide
    html_document: 
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(dplyr)
library(tidyverse)
```

## Data Description 

- The data is from the FEMA website


- The summary of Public Assistance Funded Projects (https://www.fema.gov/openfema-data-page/public-assistance-funded-projects-details-v1) 
& Preliminary Damage Assessment (https://www.fema.gov/disasters/disaster-declarations?field_dv2_state_territory_tribal_value=All&field_year_value=All&field_dv2_declaration_type_value=All&field_dv2_incident_type_target_id_selective=49124)


- Interested in hurricanes being declared as disasters from year 2009 - 2018

## EDA -- Number of hurricanes being declared as disasters in each year

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(kableExtra)
funding <- read.csv("funding.csv")
# Number of hurricanes in each year
hurr <- funding %>% 
  group_by(year) %>% 
  summarize(number = length(unique(disasterNumber)))
kable(hurr)
```

## EDA -- Distribution of Different Damage Categories Among the Years

```{r, echo = FALSE, message = FALSE, warning = FALSE}
damage <- funding %>% 
  group_by(damageCategory, year) %>% 
  summarize(number = length(damageCategory))
ggplot(damage) +
      geom_col(aes(x = damageCategory, y = number, fill = factor(year))) +
      ggtitle("Number of Projects for Different Damage Category") +
      labs(x = "Damage Category", y = "Number of Projects", fill = "Year") + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      theme(plot.title = element_text(hjust = 0.5))
```

## Total Project Amount (in dollars) from 2009 to 2018

```{r, echo = FALSE, message = FALSE}
library(maps)
state <- map_data("state")
funding_sum <- read.csv("funding_sum.csv")
ggplot() + geom_polygon(data = funding_sum, aes(x = long, y = lat, group = group,
                                                fill = range), 
                        color = "grey", size = 0.2, alpha = 1.6) + 
  geom_polygon(data = state, aes(x = long, y = lat, group = group),
                 color = "black", fill = "white", size = 0.2, alpha = 0.3) +
      scale_fill_brewer(palette = "Blues") +
  ggtitle("Project Amount from 2009 to 2018") +
  # Center the title
  theme(plot.title = element_text(hjust = 0.5))
```

## Interactive map with plotly

```{r, message = FALSE, echo = FALSE, warning = FALSE}
library(plotly)
funding_year <- read.csv("funding_year.csv")
# Create hover on the map
funding_year$hover <- with(funding_year, paste("State:", state,"<br>","Project Number:",number,"<br>")) 
project_number <- plot_geo(funding_year, locationmode = 'USA-states') 
project_number <- project_number %>% add_trace(
  locations = ~stateCode,
  type = 'choropleth',
  z = ~number,
  text = ~hover,
  colorscale = "Blues"
)
# Add title
project_number <- project_number %>% layout(
  title = 'Number of Hurricane Project from 2009 - 2018'
)

project_number
```

## Statewide per capita impact (in dollars)

```{r, message = FALSE, echo = FALSE, warning = FALSE}
# Load the data in 
assistance <- read.csv("FEMA.csv")
ggplot(data = assistance, aes(Year, Statewide_per_capita_impact))+
  geom_point(aes(color = State), alpha=0.5)+
  labs(title = "Year vs Per Capita Assistance", 
       y = "Statewide per capita impact ($)", x = "Year")+
  theme(plot.title = element_text(size = 20),
        axis.text = element_text(size =  14),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))
```


## Thank you!!